Best Practices for Single Alchemical Free Energy Calculations [Article v\versionnumber]

### 2025-06-08 paper/manuscript.tex
comments on weak binders section

% aux. comments Stefan Boresch -> Stefan Boresch
\newcommand{\sbnote}[1]{%
  {\bfseries{}[SB: }%
  {\textcolor{blue}{#1}}{\bfseries{}]}
  }
\newcommand{\rref}{{\bfseries[REFs]}{}}

@@

In the statistical mechanics formulation of Gilson et al.~\cite{gilson1997statisticalthermodynamic}, described in Section \ref{sec:non-covalent-molecular-binding}, the standard state concentration appears explicitly in Eq.~(\ref{eq:DG0-binding-constant-def}) and the binding pose is defined by the indicator function $I(\zeta)$ that characterizes the partition function of the complex, Eq.~(\ref{eq:intra-zRL-def}).  The role of the standard state and binding site definitions are particularly evident when expressing Eq.~(\ref{eq:DG0-binding-constant-def}) in terms of an ensemble average. To do so, we express the integrals in the numerator and denominator of Eq.~(\ref{eq:DG0-binding-constant-def}) over the same set of degrees of freedom. Note that the intramolecular configurational partition function of the complex, $z_{RL(v)}$, includes six more degrees of freedom--the position and orientation of the ligand relative to the receptor collectively described by the variable $\zeta$--than the product of the intramolecular partition functions, $z_{R(v)} z_{L(v)}$, of the separated receptor and ligand. To correct this discrepancy, we multiply and divide the right-hand side of Eq.~(\ref{eq:DG0-binding-constant-def})  by the quantity
\begin{equation} \label{eq:Vsite-def}
    V_{\rm site} \Omega = \int d\zeta I(\zeta) \, ,
\end{equation}
%\Omega is in my opinion a better symbol of the orientational extent than \xi_L. Lowercase greek letters are usually used
%for dynamical variables rather than integrated quantities
which measures the extent of the binding pose macrostate, where $V_{\rm site}$ corresponds to the spatial extent and $\Omega$ the orientational extent. The right-hand side of Eq.~(\ref{eq:Vsite-def}) combined with the integrals $z_{R(v)}$ and $z_{L(v)}$ yields a partition function similar to Eq.~(\ref{eq:intra-zRL-def}) but corresponding to a complex in which ligand and receptor are uncoupled~\cite{gallicchio2011recent} as if they were at infinite distance apart. The result is
\begin{equation}\label{eq:DG0-binding-constant-def-average}
K_b = C^\circ V_{\rm site} \frac{\Omega}{8 \pi^2} \langle e^{-\beta \Delta \Psi}  \rangle_0 
\end{equation}
where, similarly to Eq.~(\ref{eq:zratio-solv-average}),
\begin{equation}\label{eq:pert-energy-def}
\Delta \Psi(x_R, x_L, \zeta) =  \Psi_v(x_R, x_L, \zeta) - \left[  \Psi_v(x_R) + \Psi(x_L) \right]
\end{equation}
is the solvent-averaged binding energy of the complex when receptor and ligands are in intramolecular configurations $x_R$ and $x_L$. The ligand is in position and orientation $\zeta$ relative to the receptor (the uncoupled potential energy does not depend on  the position and orientation of the ligand relative to the receptor), and $\langle \ldots \rangle_0$ represents an ensemble average in the uncoupled ensemble. Finally, note that $\Omega = 8 \pi^2$ when the binding pose indicator function does not depend on the orientation of the ligand and Eq.~(\ref{eq:DG0-binding-constant-def-average}) simplifies to
\begin{equation}\label{eq:DG0-binding-constant-def-average-noOmega}
K_b = C^\circ V_{\rm site} \langle e^{-\beta \Delta \Psi}  \rangle_0 
\end{equation}
which, expressed in free energy units, yields
\begin{eqnarray}\label{eq:DG0-def-Vsite}
  \Delta G^\circ_b &=&  - k_B T \ln K_b = - k_B T \ln C^\circ V_{\rm site} - k_B T \ln \langle e^{-\beta \Delta \Psi}  \rangle_0 \nonumber \\
{\rm \ }  &=& \Delta G^\circ_{b,{\rm ideal}} + \Delta G_{b,{\rm exc.}}
\end{eqnarray}

The term $\Delta G^\circ_{b,{\rm ideal}} = - k_B T \ln C^\circ V_{\rm site}$ in Eq.~(\ref{eq:DG0-def-Vsite}) is interpreted as the \emph{ideal} component of the standard binding free energy, that is the binding free energy that would be measured in the absence of ligand-receptor interactions ($\Delta \Psi = 0$) and when the ligand's concentration is $C^\circ$. The ideal term is responsible for the dependence of the standard binding free energy on the standard state concentration. The arguably inappropriate term ``standard state correction'' is sometimes used in the literature to refer to $\Delta G^\circ_{b,{\rm ideal}}$. As seen above, the ideal term is a fundamental aspect originating from the statistical mechanics theory of molecular binding rather than a mere correction to an inexact formulation.

@@

As discussed above, however, the situation for binding is more complex and requires special care. The first important aspect is that binding free energy calculations must be performed with binding site restraints that implement the binding site indicator function $I(\zeta)$ by forcing the ligand to occupy the binding site even when it is decoupled from the receptor. Doing otherwise would fail to specify the bound macrostate of the complex for which we seek the free energy. Alchemical binding free energy simulations where the ligand leaves the binding site are not set up correctly and return undefined free energy estimates. Because it is a discontinuous function of the position and orientation of the ligand relative to the receptor, in molecular dynamics applications, the indicator function is often approximated by a flat-bottom restraint~\cite{chen2007can} (see below). Harmonic restraints are not as suitable in this respect because they favor a specific microstate at the bottom of the harmonic potential rather than specifying a region of conformational space.

@@

\paragraph{Several choices of binding site restraints are possible.}
A variety of restraint types are commonly used in alchemical binding free energy calculations. Before going into their characteristics, we must acknowledge an important and often unrecognized distinction between the binding site restraints corresponding to the indicator function $I(\zeta)$ necessary for the implementation of the theory and optional auxiliary restraints and biasing potentials sometimes used to speed up the convergence of free energy calculations. Binding site restraints are an integral part of the molecular complex definition and are typically maintained throughout the alchemical process. In particular, they must be present unchanged at the unbound and bound states. In contrast, unless thermodynamic reweighting is applied in post-processing, auxiliary restraints cannot be present at the endstates to avoid affecting their free energies. Hence, they are typically turned on during an initial stage, starting from the uncoupled state, and are turned off in the final stage, near the bound state. In the alchemical binding free energy literature, often harmonic restraint potentials imposed in an initial stage and removed in a final stage are described as implementing standard state terms. However, because they utilize harmonic potentials that cannot correctly describe the binding pose macrostate (see above) and because they are turned on and then off, these are actually auxiliary restraints introduced to enhance convergence without affecting, in principle, binding free energy estimates.

@@

Recently, more restrictive restraint schemes have been proposed that avoid relying on a few critical reference atoms. Instead, these aim to simplify restraint selection by using larger sets of atomic positions in the receptor and ligand. Fu et al. suggested restraining the six relative rigid-body degrees of freedom derived by finding the rotation of the ligand, which minimises the RMSD to a reference structure of the receptor-ligand complex (after accounting for rotation and translation of the receptor)~\cite{fu2017new}. Salari et al. proposed restraining the ``distance-from-bound-configuration'' (DBC) variable, which is the RMSD of ligand coordinates in the frame of reference of the binding site~\cite{salari2018streamlined, ebrahimi2022symmetry}. DBC may allow the binding pose to be closely preserved as the ligand intermolecular interactions are removed, but has the disadvantage of coupling the internal and relative external degrees of freedom of the receptor and ligand, making them unsuitable for describing the binding site macrostate and the corresponding standard state terms. However, these restraints are practical as auxiliary restraints where additional steps are used to introduce and release the DBC restraints. These restraint schemes are implemented in open-source workflows~\cite{fu2021bfee2, fu2022accurate, santiago-mcrae2023computing}.

As discussed above, binding site restraints are commonly implemented using quadratic flat-bottom restraints, which impose a restoring harmonic  force similar to harmonic potentials but only if the ligand leaves a specific region~\cite{chen2007can}. Often, binding site restraints do not limit the rotational degrees of freedom of the ligand, and the corresponding standard state term is zero; see Eq.~(\ref{eq:DG0-def-Vsite}).

Auxiliary restraints take a variety of forms, such as simple harmonic distance restraints between the ligand and the protein~\cite{mobley2006use} as well as flat-bottom restraints. The choice of degrees of freedom and restraining potentials for auxiliary restraints is not as restrictive as for those for binding site restraints. For example, the overall ligand RMSD has been used~\cite{woo2005calculation} for this purpose. Some auxiliary restraints are designed to reduce the amount of conformational sampling at intermediate $\vec{\lambda}$ values. However, more computational effort is required to compute the free energy when the restraints are turned on and off. Additionally, such restraints can limit the exploration of alternative binding modes. This  restriction may be undesirable when using Hamiltonian $\lambda$ exchange or expanded ensemble techniques where allowing the ligand to exchange binding modes when it is non-interacting could provide sampling benefits~\cite{wang2013identifying}.
More specifically, flat-bottom restraints allow a ligand to explore multiple binding sites, and harmonic restraints allow the exploration of multiple binding modes within a site, while Boresch restraints only allow a single binding mode within a single site.
See additional discussion of the possibility of multiple binding modes in Sec.~\ref{sec:multiple_binding_modes} below.

\subsection{Absolute and relative calculations deal with some of the same issues}
\subsubsection{Handling weak binders and high dissociation rates}\label{sec:weak-binders}
In binding free energy calculations, only the conformations in which the receptor and ligand form a bound complex should be sampled from the bound states (Sec.~\ref{sec:theory}). Determining what the bound states actually are can be challenging for weakly bound ligands. For tightly bound ligands, virtually all reasonable definitions of the bound state will lead to be equivalent free energies, since the partition function will be dominated by a relatively small number of low-energy poses.

@@

In absolute calculations, this need to explicitly define a binding site applies to the fully interacting state in the complex leg of the thermodynamic cycle (top-right state in Fig.~\ref{fig:fig_absolute_thermodynamic_cycle}), while in relative calculations the binding site must be defined at both end states of the complex leg (top- and bottom-right states in Fig.~\ref{fig:fig_binding_thermodynamic_cycle}).
In principle, this requires defining which conformations are considered to be bound before running the calculation, but it is common practice to start the simulation with the ligand already placed in the binding site and rely on kinetic trapping to maintain the bound complex.

However, this strategy of using kinetic trapping to maintain the bound complex can fail when the dissociation rate of the ligand has the same or smaller order of magnitude than the length of the simulation.
This is typical of weak binders such as fragments binding shallow pockets with $\mu$M-mM affinities~\cite{georgiou2017pushing,pan2017quantitative}.
In the case of weak binders, using a flat-bottom or harmonic restraint between receptor and ligand in the bound state(s) can prevent dissociations~\cite{georgiou2017pushing,rizzi2019sampl6}.
We stress that this type of restraint is normally avoided as it generally introduces bias in the free energy estimate, which is why the restraint is usually activated only in the intermediate states in absolute calculations.
The bias can be corrected through reweighting schemes~\cite{rizzi2019sampl6}, but this post-processing step can be avoided if a flat-bottom restraint is used and the ligand is never sampled while hitting the potential wall during the simulation in the bound state as the numerical correction will be exactly zero.
It is important to note that the spring constant and/or radius parameters of the restraint effectively determine which conformations are considered to be bound.
As a consequence, these parameters must be tuned to the system so that only the binding site is accessible to the ligand.
Again, this step is particularly important for weak binders as their free energy of binding is known to be more sensitive to the definition of the binding site~\cite{gilson1997statisticalthermodynamic}.

In absolute calculations, this restraint can substitute or be added to the restraint used to handle the standard state correction (Sec.~\ref{sec:standardstate-restraints}).
In the latter case, however, care must be taken when computing the standard state correction.
When multiple restraining potentials are active in the non-interacting state, the correction can generally be computed only through numerical integration.
Alternatively, one can adopt a protocol that removes the bound-state restraint in the non-interacting state.
Finally, even for tight binders, dissociation events can be enhanced by methods such as Hamiltonian replica exchange~\cite{sugita2000multidimensional,chodera2011replica,wang2013identifying} and expanded ensemble~\cite{lyubartsev1992new,li2007simulated}, especially in absolute free energy calculations using harmonic or flat-bottom restraints.
In the latter case, dissociations can be averted simply by increasing the spring constant and/or reducing the radius of the restraint potential to prevent the exploration of ligand conformations outside the binding site in the decoupled state (bottom-right state in Fig.~\ref{fig:fig_absolute_thermodynamic_cycle}) that could be propagated to the bound state.

>>>>

% aux. comments Stefan Boresch -> Stefan Boresch
\newcommand{\sbnote}[1]{%
  {\bfseries{}[SB: }%
  {\textcolor{blue}{#1}}{\bfseries{}]}
}

% aux. comments Emilio Gallicchio
\newcommand{\egnote}[1]{%
  {\bfseries{}[EG: }%
  {\textcolor{green}{#1}}{\bfseries{}]}
}

\newcommand{\rref}{{\bfseries[REFs]}{}}

@@

In the statistical mechanics formulation of Gilson et al.~\cite{gilson1997statisticalthermodynamic}, described in Section \ref{sec:non-covalent-molecular-binding}, the standard state concentration appears explicitly in Eq.~(\ref{eq:DG0-binding-constant-def}) and the binding pose is defined by the indicator function $I(\zeta)$ that characterizes the partition function of the complex; Eq.~(\ref{eq:intra-zRL-def}).  The role of the standard state and binding site definitions are particularly evident when expressing Eq.~(\ref{eq:DG0-binding-constant-def}) in terms of an ensemble average. To do so, we express the integrals in the numerator and denominator of Eq.~(\ref{eq:DG0-binding-constant-def}) over the same set of degrees of freedom. Note that the intramolecular configurational partition function of the complex, $z_{RL(v)}$, includes six more degrees of freedom--the position and orientation of the ligand relative to the receptor collectively described by the variable $\zeta$--than the product of the intramolecular partition functions, $z_{R(v)} z_{L(v)}$, of the separated receptor and ligand. To correct this discrepancy, we multiply and divide the right-hand side of Eq.~(\ref{eq:DG0-binding-constant-def})  by the quantity
\begin{equation} \label{eq:Vsite-def}
    V_{\rm site} \Omega = \int d\zeta I(\zeta) \, ,
\end{equation}
%\Omega is in my opinion a better symbol of the orientational extent than \xi_L. Lowercase greek letters are usually used
%for dynamical variables rather than integrated quantities
which measures the extent of the binding pose macrostate, where $V_{\rm site}$ corresponds to the spatial extent and $\Omega$ the orientational extent. The right-hand side of Eq.~(\ref{eq:Vsite-def}) combined with the integrals $z_{R(v)}$ and $z_{L(v)}$ yields a partition function similar to Eq.~(\ref{eq:intra-zRL-def}) but corresponding to a complex in which ligand and receptor are uncoupled~\cite{gallicchio2011recent} as if they were at infinite distance apart. The result is
\begin{equation}\label{eq:DG0-binding-constant-def-average}
K_b = C^\circ V_{\rm site} \frac{\Omega}{8 \pi^2} \langle e^{-\beta \Delta \Psi}  \rangle_0 
\end{equation}
where, similarly to Eq.~(\ref{eq:zratio-solv-average}),
\begin{equation}\label{eq:pert-energy-def}
\Delta \Psi(x_R, x_L, \zeta) =  \Psi_v(x_R, x_L, \zeta) - \left[  \Psi_v(x_R) + \Psi(x_L) \right]
\end{equation}
is the solvent-averaged binding energy of the complex when receptor and ligands are in intramolecular configurations $x_R$ and $x_L$, the ligand is in position and orientation $\zeta$ relative to the receptor (the uncoupled potential energy does not depend on  the position and orientation of the ligand relative to the receptor), and $\langle \ldots \rangle_0$ represents an ensemble average in the uncoupled ensemble. Finally, note that $\Omega = 8 \pi^2$ when the binding pose indicator function does not depend on the orientation of the ligand and Eq.~(\ref{eq:DG0-binding-constant-def-average}) simplifies to
\begin{equation}\label{eq:DG0-binding-constant-def-average-noOmega}
K_b = C^\circ V_{\rm site} \langle e^{-\beta \Delta \Psi}  \rangle_0 
\end{equation}
which, expressed in free energy units, yields
\begin{eqnarray}\label{eq:DG0-def-Vsite}
  \Delta G^\circ_b &=&  - k_B T \ln K_b = - k_B T \ln C^\circ V_{\rm site} - k_B T \ln \langle e^{-\beta \Delta \Psi}  \rangle_0 \nonumber \\
{\rm \ }  &=& \Delta G^\circ_{b,{\rm ideal}} + \Delta G_{b,{\rm exc.}}
\end{eqnarray}

The term $\Delta G^\circ_{b,{\rm ideal}} = - k_B T \ln C^\circ V_{\rm site}$ in Eq.~(\ref{eq:DG0-def-Vsite}) is interpreted as the \emph{ideal} component of the standard binding free energy, that is the binding free energy that would be measured in the absence of ligand-receptor interactions ($\Delta \Psi = 0$) and when the ligand's concentration is $C^\circ$. The ideal term is responsible for the dependence of the standard binding free energy on the standard state concentration. The arguably inappropriate term ``standard state correction'' is sometimes used in the literature to refer to $\Delta G^\circ_{b,{\rm ideal}}$. As seen above, the ideal term is a fundamental aspect of the statistical mechanics theory of molecular binding rather than a mere correction to an inexact formulation.

@@

As discussed above, the situation for binding is more complex and requires special care. The first important aspect is that binding free energy calculations must be performed with binding site restraints that implement the binding site indicator function $I(\zeta)$ by forcing the ligand to occupy the binding site even when it is decoupled from the receptor. Doing otherwise would fail to specify the bound macrostate of the complex for which we seek the free energy. Moreover, the indicator function prescribes that the ligand remains in the binding site region throughout the alchemical process, including when it is uncoupled from the receptor. Consequently, alchemical binding free energy simulations where the ligand leaves the binding site are not set up correctly and return undefined free energy estimates. Because it is a discontinuous function of the position and orientation of the ligand relative to the receptor, molecular dynamics implementations often approximate the indicator function by means of flat-bottom restraints~\cite{chen2007can} (see below). Harmonic restraints are not as suitable in this respect because they favor a specific microstate at the bottom of the harmonic potential rather than specifying a region of conformational space.

@@

\paragraph{Several choices of binding site restraints are possible.}
A variety of restraint types are commonly used in alchemical binding free energy calculations. Before going into their characteristics, we must acknowledge an important and often unrecognized distinction between the binding site restraints corresponding to the indicator function $I(\zeta)$ necessary for the implementation of the theory and optional auxiliary restraints and biasing potentials sometimes used to speed up the convergence of free energy calculations. Binding site restraints are an integral part of the molecular complex definition and are typically maintained throughout the alchemical process. In particular, they must be present unchanged at the unbound and bound states. In contrast, unless thermodynamic reweighting is applied in post-processing, auxiliary restraints cannot be present at the endstates to avoid affecting their free energies. Hence, they are typically turned on during an initial stage, starting from the uncoupled state, and are turned off in the final stage, near the bound state. In the alchemical binding free energy literature, often harmonic restraint potentials imposed in an initial stage and removed in a final stage are described as implementing binding sites restraints. However, because they utilize harmonic potentials that cannot correctly describe the binding pose macrostate (see above) and because they are turned on and then off, these are instead auxiliary restraints introduced to enhance convergence without affecting, in principle, binding free energy estimates.

@@

Recently, degrees of freedom for binding site and auxiliary potentials based on larger sets of atomic positions of the receptor and ligand have been put forward.
Fu et al. suggested restraining the six relative rigid-body degrees of freedom derived by finding the rotation of the ligand, which minimises the RMSD to a reference structure of the receptor-ligand complex (after accounting for rotation and translation of the receptor)~\cite{fu2017new}. Salari et al. proposed restraining the ``distance-from-bound-configuration'' (DBC) variable, which is the RMSD of ligand coordinates in the frame of reference of the binding site~\cite{salari2018streamlined, ebrahimi2022symmetry}. DBC may allow the binding pose to be closely preserved as the ligand intermolecular interactions are removed, but has the disadvantage of coupling the internal and relative external degrees of freedom of the receptor and ligand, making them unsuitable for describing the binding site macrostate and the corresponding standard state terms. However, these restraints are practical as auxiliary restraints where additional steps are used to introduce and release the DBC restraints. These restraint schemes are implemented in open-source workflows~\cite{fu2021bfee2, fu2022accurate, santiago-mcrae2023computing}.

As discussed above, binding site restraints are commonly implemented using quadratic flat-bottom restraints, which impose a restoring harmonic  force similar to harmonic potentials but only if the ligand leaves a specific region~\cite{chen2007can}. Often, binding site restraints do not limit the rotational degrees of freedom of the ligand, and the corresponding standard state term is zero; see Eq.~(\ref{eq:DG0-def-Vsite}). However, it might be challenging to fully explore the bound macrostate at all alchemical states when it is only limited by receptor-ligand positional restraints. As further discussed below, auxiliary restraints are particularly helpful in the latter circumstance to limit the extent of conformational sampling necessary to achieve rapid convergence of binding free energy estimates.

Auxiliary restraints take a variety of forms, such as simple harmonic distance restraints between the ligand and the protein~\cite{mobley2006use} as well as flat-bottom restraints. The choice of degrees of freedom and restraining potentials for auxiliary restraints is not as restrictive as for those for binding site restraints. For example, the overall ligand RMSD has been used~\cite{woo2005calculation} for this purpose. Auxiliary restraints are designed to reduce the amount of conformational sampling at intermediate $\vec{\lambda}$ values. However, more computational effort is required to compute the free energy when the restraints are turned on and off. Additionally, such restraints can limit the exploration of alternative binding modes. This  restriction may be undesirable when using Hamiltonian $\lambda$ exchange or expanded ensemble techniques where allowing the ligand to exchange binding modes when it is non-interacting could provide sampling benefits~\cite{wang2013identifying}.
More specifically, flat-bottom restraints allow a ligand to explore multiple binding sites, and harmonic restraints allow the exploration of multiple binding modes within a site, while Boresch restraints only allow a single binding mode within a single site.
See additional discussion of the possibility of multiple binding modes in Sec.~\ref{sec:multiple_binding_modes} below.

\subsection{Absolute and relative calculations deal with some of the same issues} 
\subsubsection{Handling weak binders and high dissociation rates}\label{sec:weak-binders}
\egnote{Is this section necessary? We could add a comment regarding weak binders in the preceding section clarifying how the theory covers weak binders (as well as non-binders) and the precautions the user should take to obtain accurate estimates (relative to the experiments).}
In binding free energy calculations, only the conformations in which the receptor and ligand form a bound complex should be sampled from the bound states (Sec.~\ref{sec:theory}). Determining what the bound states actually are can be challenging for weakly bound ligands. For tightly bound ligands, virtually all reasonable definitions of the bound state will lead to be equivalent free energies, since the partition function will be dominated by a relatively small number of low-energy poses.

@@

In absolute calculations, the need to explicitly define a binding site applies to the fully interacting state in the complex leg of the thermodynamic cycle (top-right state in Fig.~\ref{fig:fig_absolute_thermodynamic_cycle}),  while in relative calculations the binding site must be defined at both end states of the complex leg (top- and bottom-right states in Fig.~\ref{fig:fig_binding_thermodynamic_cycle}). \egnote{I do not agree with the latter. Binding site restraints apply to the coupled and uncoupled states. The same is true for RBFE for both ligands. (see above)}
In principle, this requires defining which conformations are considered to be bound before running the calculation, but it is common practice to start the simulation with the ligand already placed in the binding site and rely on kinetic trapping to maintain the bound complex.

However, this strategy of using kinetic trapping to maintain the bound complex can fail when the dissociation rate of the ligand has the same or smaller order of magnitude than the length of the simulation.
This is typical of weak binders such as fragments binding shallow pockets with $\mu$M-mM affinities~\cite{georgiou2017pushing,pan2017quantitative}.
In the case of weak binders, \egnote{I would say for all ligands: strong, weak, as well as non-binders.} using a flat-bottom or harmonic restraint \egnote{Harmonic restraints are not suitable as binding site restraints, see above} between receptor and ligand in the bound state(s) can prevent dissociations~\cite{georgiou2017pushing,rizzi2019sampl6}.
We stress that this type of restraint is normally avoided as it generally introduces bias in the free energy estimate, which is why the restraint is usually activated only in the intermediate states in absolute calculations. \egnote{The latter statement is incorrect. Binding site restraints are always necessary in principle. }
The bias can be corrected through reweighting schemes~\cite{rizzi2019sampl6}, \egnote{Auxiliary restraints and biasing potentials require reweighting. But binding site restraints do not generally require any action in post-processing except for adding the $C^\circ V_{\rm site}$ term. } but this post-processing step can be avoided if a flat-bottom restraint is used and the ligand is never sampled while hitting the potential wall during the simulation in the bound state as the numerical correction will be exactly zero. \egnote{The ligand frequently hitting the boundary of the flat-bottom restraints while in the coupled state is indicative of an inappropriate definition of the binding site: the ligand is attempting to access a lower free energy binding pose. The free energy estimate is unlikely to agree with experiments but the calculation is otherwise valid. It reports the binding free energy in the binding pose specified by the region allowed by the flat-bottom binding site restraint.}
It is important to note that the spring constant and/or radius parameters of the restraint effectively determine which conformations are considered to be bound.
As a consequence, these parameters must be tuned to the system so that only the binding site is accessible to the ligand.
Again, this step is particularly important for weak binders \egnote{but necessary, in principle, for all binders} as their free energy of binding is known to be more sensitive to the definition of the binding site~\cite{gilson1997statisticalthermodynamic}.

In absolute calculations, this restraint can substitute or be added to the restraint used to handle the standard state correction (Sec.~\ref{sec:standardstate-restraints}).
In the latter case, however, care must be taken when computing the standard state correction.
When multiple restraining potentials are active in the non-interacting state, the correction can generally be computed only through numerical integration.
Alternatively, one can adopt a protocol that removes the bound-state restraint in the non-interacting state.
Finally, even for tight binders, dissociation events can be enhanced by methods such as Hamiltonian replica exchange~\cite{sugita2000multidimensional,chodera2011replica,wang2013identifying} and expanded ensemble~\cite{lyubartsev1992new,li2007simulated}, especially in absolute free energy calculations using harmonic or flat-bottom restraints.
In the latter case, dissociations can be averted simply by increasing the spring constant and/or reducing the radius of the restraint potential to prevent the exploration of ligand conformations outside the binding site in the decoupled state (bottom-right state in Fig.~\ref{fig:fig_absolute_thermodynamic_cycle}) that could be propagated to the bound state. \egnote{I disagree. Binding site restraints are never turned off, especially at the uncoupled state, which would result in the ligand wandering away from the receptor. Additional restraints can be applied to reduce the extent of conformational sampling in the intermediate alchemical steps. Conversely, energy barrier-flattening potentials can be applied to accelerate conformational sampling. These auxiliary potentials must be turned on and then turned off so that they are not active at the endstates. Alternatively, thermodynamic reweighting must be applied in post-processing to remove their effects on the binding free energy estimate. }
